---
title: "task11"
author: "Milda Zarankaite"
date: "Monday, April 18, 2016"
output: html_document
---

Lab Session 4. Task 1
=====
Tasks reference - @Hyndman2014a

Library fpp recomended for all of the tasks :
```{r, message=FALSE}
library(fpp)
```

##Task:
####1. Use ets() to find the best ETS model for the price of eggs (data set eggs). How does this model compare to the one you found in the previous lab session?

1.Let's look at the data
```{r}
plot(eggs, main="Data set Eggs")
```

2.Finding the best ETS model while using ets().
```{r}
mod1<-ets(eggs)
mod1f<-forecast(mod1, h=100)
plot(mod1f)

mod2<-ets(eggs, model="MMN") #exponential
mod2f<-forecast(mod2, h=100)
plot(mod2f)
```

The data is not seasonal so choosed models has no seasonal trend. The 1st model suggested by ets function is MNN. 2nd model MMN is using exponential smoothing.

Model, created  in task 10 from the same data and chosed as the best:
```{r}
mod3 <- holt(eggs, h=100, exponential=TRUE, beta=0.001)
plot(mod3)
```

3. Compairing these 3 models with accuracy and AIC  
* Checking models accuracy:
```{r}
accuracy(mod1)
accuracy(mod2)
accuracy(mod3)
```
According to the accuracy, differences between models are not significant. But mod2 has the least RMSE.   
Let's check models AIC:  
(mod3 AIC we get from:
```{r, message=FALSE}
mod3[1]
AICmod3<-1042.862
```
)
```{r, echo=FALSE}
mod.pav<-c("mod1", "mod2", "mod3")
aic<-c(AIC(mod1),AIC(mod2),AICmod3) 
tab<- data.frame(mod.pav,aic) 
colnames(tab)<-c("Modelis", "AIC") 
tab
```

According to AIC, mod1 has the least AIC but it does not seem very important, comparing to others. Moreover, it is known that est function chooses the best model comparing their AIC.  

As a result, I would prefer mod2 for the forecast as it has the best accuracy value and it takes into account decreasing tendency, confidence intervals are not so wide.